\begin{abstract}
    These are course notes for MAT 5190.
\end{abstract}

\section{September 8, 2021}
\subsection{Set Theory}

\begin{definition}[Sample Space]
The set of all possible outcome of an experiment, $S$ is called the \vocab{sample space} for the experiment.
\end{definition}
Note how the sample space entirely depends on the experiment. Once the sample space has been defined, we can consider \textit{collections} of possible outcomes of the experiment.
\begin{definition}[Event]
An \vocab{event} is a subset of the sample space or equivalently
an event is a collection of outcomes of an experiment.
\end{definition}
We say event $A \subseteq S$ occurs if the outcome of the experiment is in the set $A$.
\textbf{Union}: 
$A\cup B = 
\{
x: x \in A \text{ or } x \in B
\}
$
\textbf{Complement}:
$A^c = 
\{
x: x \notin A
\}$.
\textbf{Intersection}:
$
A \cap B: 
\{
x: x \in A \text{ and } x \in B
\}
$

\begin{definition}[Disjoint or mutually exclusive, pairwise disjoint]
Two events $A$ and $B$ are \vocab{disjoint (or mutually exclusive)}  if $A \cap B = \emptyset$. The events $A_1, A_2, ...$ are pairwise disjoint if $A_i \cap A_j = \emptyset$ for all $i \neq j$.
\end{definition}

\begin{example}
    Disjoint sets are sets with no points in common. Basic example is the collection
    $$
    A_i = [i, i+1), i = 0,1,2,..
    $$
    Note the $\cup_{i=1}^{\infty}A_i = [0, \infty)$
\end{example}

\begin{definition}[Partition]
If $A_1, A_2,...$ are pairwise disjoint and $ \cup_{i=1}^{\infty}A_i = S$, then the collection $A_1, A_2,...$ forms a \vocab{partition} of $S$.
\end{definition}
\subsection{Basics of Probability Theory}
\begin{definition}[Sigma algebra]
    A collection of subsets of $S$ is called a \vocab{sigma algebra} ( or \vocab{Borel field)}, denoted by $\SB$, if it satisfies the following three properties:
    
    \begin{enumerate}[(a)]
        \item $\emptyset \in \SB$ (the empty set is an element).
        \item If $A \in \SB$, then $A^c \in \SB$ (closed under complement) 
        \item If $A_1, A_2, ... \in \SB$, then, $\cup_{i=1}^{\infty}A_i \in \SB$ (closed under countable unions).
    \end{enumerate}
\end{definition}
Many different sigma algebras can be associated with a sample space $S$, the collection $\{\emptyset, S\}$ is the trivial sigma algebra. The only sigma algebra that we will be concerned with is smallest one that contains all open sets in a given sample space.

\begin{definition}[Probability Function]
    Given a sample space $S$ and a sigma algebra $\SB$, a \vocab{probability function} is a function $\PP$ with domain $\SB$ that satisfies
    \begin{enumerate}
        \item $\PP(A) \geq 0$ for all $A \in \SB$
        \item $\PP(S) = 1$
        \item If $A_1, A_2,... \in \SB$ are pairwise disjoint, then $\PP(\cup_{i=1}^{\infty}A_i) = \sum_{i=1}^{\infty}\PP(A_i)$
    \end{enumerate}
\end{definition}


\subsection{Conditional Probability and Independence}
\begin{definition}[Conditional Probability]
    If $A$ and $B$ are events in S, and $\PP(B) \geq 0$, then the \vocab{conditional} probability of $A$ given $B$ is
    $$
    \PP(A|B) = \frac{\PP(A \cap B)}{\PP(B)}
    $$
\end{definition}
Note that in the calculation of conditional probability the event $B$ becomes the sample space. The intuition is that our original sample space $S$ has been updated to $B$. All further occurrences are then calibrated with respect to their relation with $B$.

\begin{theorem}[Bayes' Rule]

$A_1, A_2,...$ be a partition of the sample space, and let $B$ be any set. The for each $i = 1,2,...$

$$
\PP(A_i|B) = \frac{\PP(B|A_i)\PP(A_i)}{\sum_{j=1}^{\infty}\PP(B|A_j)\PP(A_j)}
$$
\end{theorem}

\begin{definition}[Independent, Mutually independent]
    Two events $A$ and $B$ are \vocab{statistically independent} if
    $$
    \PP(A \cap B) = \PP(A)\PP(B)
    $$
    A collection of events $A_1,...,A_n$ are \vocab{mutually independent} if for any subcollection $A_{i1},...A_{ik}$, we have 
    $$
    \PP\left( \bigcup_{j=1}^{k}A_{ij} \right ) = \prod_{j=1}^{k} \PP(A_{ij})
    $$
\end{definition}
\subsection{More Basics}
\begin{definition}[Random variable]
    A \vocab{random variable} is a function from sample space $S$ into $\RR$.
\end{definition}
With every random variable $X$ we associate a function called the cumulative distribution function of $X$.
\begin{definition}[cdf, pmf, pdf]
    The \vocab{cumulative distribution function} (cdf) of a random variable $X$, denoted by $F_{X}(x)$, is defined by 
    $$
    F_X(x) = \PP_X(X\leq x), \text{ for all } x.
    $$
    The \vocab{probability mass function} (pmf) of a \textit{discrete} random variable $X$ is given by
    $$
    f_X(x) = \PP_X(X=x) \text{ for all } x.
    $$
    The \vocab{probability density function} (pdf), $f_X(x)$ of a \textit{continuous} random variable is $X$ is the function that satisfies
    $$
    F_X(x) = \int_{-\infty}^{x}f_X(t)dt \text{ for all } x
    $$
    
    
\end{definition}
